\chapter{Acknowledgments}

First of all, I want to thank {\bf you}, the reader, for picking up this thesis. Thank you for reading through all the chapters, or if you are anything like me, thank you for coming straight to the acknowledgment section. Whenever I find a Ph.D. thesis, this section is my first stop. The journey of any doctoral candidate is long with all its ups and downs, which often (hopefully) culminates in a collection of chapters. Every journey is unique, and one can understand the person on this journey by knowing who they acknowledge after completing such a taxing yet inciting endeavor. My Ph.D. journey at the Physics of Fluids (PoF) group, like all others, has been made possible by the support of many people. I want to spend the next few pages sharing my gratitude towards them. 

{\bf Detlef}, thank you so much for giving me this opportunity to work with you and all the independence to pursue the wide variety of intellectually stimulating problems in this thesis. We first got in touch via email and Skype (in the old pre-pandemic, non-zoom days). I received a one-line reply to my long request seeking a research position at PoF, which I now know to be your efficient style, as these one-line emails have become a part of my daily life. I cannot express in words how much I have enjoyed working with you over the past four years. Someone once told me that working with you is like working with Albert Einstein but in fluid dynamics. After four incredible years, I can attest to this fact. I particularly enjoy our meetings when they do not have an end time. I get to hear your thoughts and learn about the rich history of fluid dynamics, which are equally inspiring. I am often amazed by your energy. For example, I remember once, while we recorded lectures for the Advanced Fluid Dynamics course, we spent almost eight hours in your office. My job was simply to hit record and listen in as you taught the modules on turbulence, boundary layers, and instabilities. It was incredibly one of the most informative sessions of my life. Even after eight hours, you were still as energetic as when we started. Indeed, that day I witnessed the perfect example of what happens when one does what they love. In your words, the only essential thing in one's life is the brain and the passion for science; everything else is just background noise. {\bf Detlef}, for all the hours that you spent with me working out the Physics, thanks a lot. I want to continue this quest in the world of fluid dynamics by asking relevant questions, as you so often told me, and answer them to the best of my abilities. I am looking forward to working ever so closely with you in the future. I have a lot yet to learn from you, and I am sure we will have many long sessions filled with intellectually stimulating discussions in the future. 

{\bf Mazi}, the transition at the beginning of anyone's Ph.D. journey is often overwhelming and confusing. Thank you for helping me through this phase. Working on chapter six of this thesis, I learned a lot about fluid dynamics and non-Newtonian flows. Furthermore, you have always lent an ear whenever I felt stuck in the past four years. The long scientific discussions and the ideas you passed on have been invaluable to me. But, of course, your influence on this thesis is not restricted to just being the go-to person for quenching Scientific curiosity. You have also acted as the perfect Yoda on this Ph.D. journey, both academically and otherwise. Thank you so much for inspiring me all through the process. I look forward to working together on many ideas and projects. 

I have been fortunate to be always surrounded by some of the most incredible mentors, {\bf Detlef}, {\bf Mazi}, {\bf St{\'e}phane Z.}, {\bf Andrea}, {\bf John S.}, {\bf Jean-Philippe}, {\bf Micka{\"e}l}, and {\bf Arup}. The privilege often reminds me of these words taken from a letter Newton wrote to Robert Hooke, \lq\lq If I have seen further, it is by standing on the shoulders of giants.\rq\rq\, I am indebted to all of you for encouraging and showing me the way. {\bf St{\'e}phane Z.}, thank you for teaching me all I know about the volume of fluid method. Given that this method forms the backbone of my entire Ph.D. thesis, it is safe to say that this thesis would not exist without the GIAN course on two-phase flow at the Indian Institute of Technology Roorkee. You are always eager to discuss science whenever we meet during conferences or during your visits to PoF. These discussions have benefited me tremendously throughout my career. {\bf Andrea}, your enthusiasm, passion, and search for new fluid dynamics problems have motivated me. The long sessions with you always leave me full of thoughts and ideas. {\bf Arup}, I would also like to thank you for introducing me to the fascinating world of fluid dynamics. The journey that started in your office about eight years ago has now reached this stage, with hopefully more to come. 

Here, I would like to take the opportunity to thank my {\bf committee members}. Thanks a lot for agreeing to be a part of my graduation committee and for spending so much of your precious time reading this thesis. I look forward to interacting with you all on the day of the defense and in the future. 

Next, I want to thank my paranymphs, {\bf UJ} and {\bf Udo}. First of all, thanks for agreeing to accompany me on the stage as I defend this Ph.D. thesis and for dressing up formally for the occasion. I know how much both of you share my discomfort in adhering to formal attire. {\bf Udo}, your passion for perfection is a force to reckon with. Thanks for always being critical of my work and providing honest feedback. We had a great time working on the Taylor-Culick retractions, which formed the largest chapter of this thesis. I hope we will continue our collaboration and the long walks that consistently contributed to pushing these collaborations. I hope we will tackle many more interesting fluid dynamics problems in the future. {\bf UJ}, I have found a great friend in you. We have spent many weekends in our offices discussing all things related to fluid dynamics and life. {\bf Yanshen}, {\bf Peter}, you, and I made the weekend warriors at PoF. We also had a wonderful and insightful time in Copenhagen. I have also always enjoyed our long walks. By the way, {\bf UJ}, we have known each other since my first visit to PoF when we went out for dinner following my interview with {\bf Detlef} where he offered me this doctoral position. Little did I know that we share a common friend, who, as it turns out, is one of my closest allies in life. Thank you, {\bf Sonali} for always being available at a phone call's notice and for all the help you have provided throughout these years.

{\bf Pierre}, we have shared an office for almost three years. We have interacted on several open projects and ideas throughout this time and often had heated but spirited discussions about our works. Collaborating with you on chapters 2 and 3 of this thesis has been excellent. I have thoroughly enjoyed our brainstorming sessions. In one of them, we also came up with the title of this thesis. In fact, the first part of this work on drop impact processes draws motivation from these discussions and your thesis. The latter has also contributed to improving my understanding of the French language. I also want to thank you for always lending an ear to my, at times, crazy ideas. 

Next, I would like to thank all my collaborators and coauthors, {\bf Claas-Willem}, {\bf Olinka}, {\bf Doris}, {\bf Jon}, {\bf Bin}, {\bf Cunjing}, {\bf Srinath}, and {\bf Pallav}. I have learned a lot from all of you and am looking forward to several stimulating collaborations in the future. 

Next, I want to thank the entire group at PoF, my professional family. The Wednesday seminars are a tour de force. We get to share ideas, listen to fluid dynamics leaders from across the globe, and most importantly, enjoy stimulating discussions. Being a part of this group is a privilege I have enjoyed as a young researcher. I want to thank {\bf Joanita} for her superhuman organizational skills. You show an ideal combination of speed and precision. Thank you for always having a solution to any and all of my problems. 

Of course, the group is only as good as the people in it. At PoF, I interacted with amazing fluid dynamicists. Where else would one find a fluid dynamics rock star like {\bf Leen} whose contributions to fluid dynamics research and several lives are both incredible and humbling. 

{\bf Devaraj}, thank you for agreeing to be the chairperson of my committee and for all the positiveness during our interactions. Your knowledge of all matters, Physics or otherwise, is terrific. I got a first-hand experience during the pop quiz at Landal Village. It was a pleasure to be a part of the team centurions and compete alongside {\bf Javier} and {\bf Lor{\'e}ne}. Indeed, it was a lovely group outing. I really enjoyed the place, the science, and most of all, the company in the bungalow \#100. Thanks a lot. 

{\bf Jacco}, thank you for the insightful conversations regarding free-surface flows. We have run into each other several times at Starbucks, and I am often intrigued by how easy it is to talk to you about everything, irrespective of the complexity. I remember the one conversation we had regarding the origin of viscosity and the no-slip boundary condition in continuum mechanics. It was, in a word, inspiring. I would also like to thank {\bf Alvaro}, {\bf Richard}, {\bf Corinna}, {\bf Dominik}, {\bf Tim}, {\bf Michel}, {\bf Marjolein}, and {\bf Christian} for always asking just the right questions during group meetings, seminars, or even at dinner tables. 

{\bf Sander}, thank you for all the chocolates that helped maintain my blood sugar level over the years. I could always rely on your (not-so) secret stash whenever I needed something. Also, thank you for giving me a chance to teach some of the Advanced Fluid Mechanics modules. The best way to learn something is to teach it, and I thank you for allowing me to gauge and sharpen my knowledge through these lectures. Lastly, thank you for proofreading the Dutch summary despite your busy schedule and for your help with making the PoF-Basilisk logo, along with all the illustrator tricks you have shown me over the years. I look forward to spending more time with you and, finally, learning Mathematica. 

{\bf Diana}, we went on several long walks together through the woods, despite being scared of the caterpillars. We also had so much fun in the pre-pandemic (almost prophetic) Pandemic games. Thank you for all the fun we had together. Also, thank you for all the relatable stories from Columbia; oops, I meant Colombia, captain correct America. 

{\bf Youssef}, we share several commonalities. We started our respective doctoral works at a gap of a few months, share the same office, and perhaps most importantly, specialize in Basilisk. Thank you for always providing a bouncing board for new and old Basilisk ideas and tricks. Now, coincidentally, we are also getting married within a few months of each other. I wish you an amazing married life. 

{\bf Maaike}, thank you for all the fun and scientific discussions we have had over the years. Thank you for always listening in with so much attention and being ever so polite. I wish you the best for your bike trip across a few continents.

{\bf Carola}, I always look forward to our coffee breaks. Thank you for all the discussions and advice you have given me throughout the years. And also, thanks a lot for allowing me to \lq steal\rq\, some of your thesis template ideas. 

{\bf Jessica}, I have always enjoyed your company. Thank you so much for all the fun and, at times, serious conversations we have had over the years.

{\bf Varghese}, {\bf Shantanu}, and {\bf Vamsi}: you formed the previous generation of doctoral candidates who graduated immediately before I joined PoF. You all were also deeply involved in my hiring interviews. Thank you for sharing your experiences and showing me what a wonderful place PoF is. There are still many PoF colleagues and friends who I would like to thank for making the last four years a delight and a blast, {\bf CS}, {\bf Yibo}, {\bf Timothy}, {\bf Robert}, {\bf Anupam}, {\bf Youan}, {\bf Morgan}, {\bf Chris}, {\bf Farzan}, {\bf Ambre}, {\bf Alex}, {\bf Martin E.}, {\bf Vanshu}, and {\bf Anja}. Thank you, {\bf Minkush} for being so kind and caring for everyone. {\bf Lijun} and {\bf Yogesh}, thank you for the several dinners we have had together and never kicking {\bf Udo} or me out even if we self-invited ourselves. I also enjoyed our cricket games together with {\bf Srinidhi}. 

My old and current office-mates: {\bf Rodrigo}, {\bf Dennis B.}, {\bf Mazi}, {\bf Martin}, {\bf Marvin}, {\bf Pierre}, and {\bf Youssef}. Thanks for always maintaining such a great environment at {\bf ME 246B}. {\bf Dennis B.}, your passion for open-source programming is unparalleled. I have always enjoyed out long discussions on this topic. 

My office neighbors, the gang in {\bf ME 24/7}: {\bf Jelle}, {\bf Pieter}, {\bf Srinidhi}, and {\bf UJ}, thank you for always providing a place to cool off and swap stories. 

Next, I would like to thank the fantastic technicians and the support team at PoF. {\bf Dennis vG.}, thank you for maintaining order at PoF. {\bf Bas} and {\bf Thomas} for helping me with all the tech support and acquiring gadgets. {\bf Gert-Wim} and {\bf Martin B.}, even though I did not participate in any experimental work during my Ph.D., it was always a delight to talk to both of you whenever we met. 

During my Ph.D., I had the privilege to engage with several motivated students, {\bf Laurence},  {\bf Coen}, {\bf Mihir}, {\bf Pim}, {\bf Steven}, {\bf Thijmen}, {\bf Tom}, {\bf Toon}, and {\bf Twan}, just to name a few. I thank you all for working with me and wish you all the best for your subsequent undertakings.

{\bf St{\'e}phane P.}, you have developed and maintained one of the largest free-software programs in our community. I started my research career with Gerris and then moved to Basilisk C. I want to thank you and the entire Basilisk community for providing a well-documented partial differential equation solver. Learning the code and implementing new modules are both very easy, thanks to your efforts.

I would be amiss if I did not mention all the love and support I have received from {\bf my family members}, including my parents, sister, and in-laws. Your willingness to support me through higher studies is commendable. In a society where graduating from college and seeking a job represent the norm, I am lucky to be associated with a family encouraging higher education. Words cannot acknowledge the support I have received from all of you. Of course, my extended family includes my fantastic close friends, {\bf Akash} and {\bf Akanksha}; thank you for always being there whenever I needed you. 

{\bf Vatsalya}, thank you for providing a sanctuary whenever I needed one and for proofreading parts of this thesis despite all the technical jargon. We have been friends since high school, and we (almost) share our first names but have pursued very different and arguably complementary career paths. My love for Physics and yours for music benefited us during high-school exams. But, I have only you to thank for introducing me to my favorite human being and now, my wife. 

{\bf Anjali}, there is nothing I can say that you do not know already. We have known each other since high school. A Physics question on Kirchhoff's circuit laws brought us together, and you have been with me ever since. You have seen me both at my best and worst. Thank you for always believing in me, even when I did not. You bring a smile to my face, irrespective of the situation. This journey would have been difficult without your constant presence in my life. My warm hugs and a big thank you to you. Thank you so much for always being there. I dedicate this thesis to you, my lovely wife and the one constant in my life. 

